Novel lattice Boltzmann method for simulation of strongly shear thinning viscoelastic fluids

### 2024-01-12 NovelLBM.tex
numeric solution

	\section{FENE-P}\label{app:FENE-P}
	\todo{stability issue, fitting parameter, see fenep repo}
	\section{Analytical solution for PTT}\label{app:PTT}
	Our derivation is analogous to Oliveira\cite{oliveira1999}. To account for the added viscous stress due to $\eta_\text{s}$, equilibrium is reached at lower polymer stresses. Therefore, Oliveiras equation (7) needs to be modified as follows.
	\begin{equation}
		\tau_{xy} = \partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}
	\end{equation}
	Where the second term is added due to the presence of the solvent viscosity. This term also propagates to Oliveiras equation (9) as follows.
	\begin{align}
		\dot{\gamma} &= f\left(\frac{2\lambda}{\eta_\text{p}}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j}-\eta_\text{s}\dot{\gamma}\right)
		\intertext{This cannot be integrated as easily as without $\eta_\text{s}$, and is only solvable semi-analytically.}
		1 &= f\left(\frac{2\lambda}{\eta_\text{p}}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right)
	\end{align}
	For the exponential version of PTT $f$ is defined as follows.
	\begin{equation}
		f\left(\Tr\ma{\tau}\right) = \exp\left(\frac{\epsilon\lambda}{\eta_\text{p}}\Tr\ma{\tau}\right)
	\end{equation}
	This can be inserted in the above equation.
	\begin{align}
		1 &= \exp\left(\frac{2\epsilon\lambda^2}{\eta_\text{p}^2}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right) \\
		0 &= \frac{2\epsilon\lambda^2}{\eta_\text{p}^2}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2 - \ln\eta_\text{p} + \ln\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right)
	\end{align}
	With this, $\dot{\gamma}$ can be found for each $r$ numerically. With $\dot{\gamma} = \partial_ru$, the result can be numerically integrated to find the velocity profile. To do this, the radial distance needs to be discretized. We split it into \num{1e5} steps and interpolate between them as needed.

>>>>

	\section{FENE-P}\label{app:FENE-P}
	\todo{stability issue, fitting parameter, see fenep repo}
	\section{Analytical solution for PTT}\label{app:PTT}
	Our derivation is analogous to Oliveira\cite{oliveira1999}. To account for the added viscous stress due to $\eta_\text{s}$, equilibrium is reached at lower polymer stresses. Therefore, Oliveiras equation (7) needs to be modified as follows.
	\begin{equation}
		\tau_{xy} = \partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}
	\end{equation}
	Where the second term is added due to the presence of the solvent viscosity. This term also propagates to Oliveiras equation (9) as follows.
	\begin{align}
		\dot{\gamma} &= f\left(\frac{2\lambda}{\eta_\text{p}}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j}-\eta_\text{s}\dot{\gamma}\right)
		\intertext{This cannot be integrated as easily as without $\eta_\text{s}$, and is only solvable semi-analytically.}
		1 &= f\left(\frac{2\lambda}{\eta_\text{p}}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right)
	\end{align}
	For the exponential version of PTT $f$ is defined as follows.
	\begin{equation}
		f\left(\Tr\ma{\tau}\right) = \exp\left(\frac{\epsilon\lambda}{\eta_\text{p}}\Tr\ma{\tau}\right)
	\end{equation}
	This can be inserted in the above equation.
	\begin{align}
		1 &= \exp\left(\frac{2\epsilon\lambda^2}{\eta_\text{p}^2}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2\right)\frac{1}{\eta_\text{p}}\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right) \\
		0 &= \frac{2\epsilon\lambda^2}{\eta_\text{p}^2}\left[\partial_xp\frac{r}{2^j} - \eta_\text{s}\dot{\gamma}\right]^2 - \ln\eta_\text{p} + \ln\left(\partial_xp\frac{r}{2^j\dot{\gamma}}-\eta_\text{s}\right) \\
		0 &= 2\epsilon\lambda^2\dot{\gamma}^2\left[\partial_xp\frac{r}{2^j\dot{\gamma}\eta_\text{p}} - \frac{\eta_\text{s}}{\eta_\text{p}}\right]^2 + \ln\left(\partial_xp\frac{r}{2^j\dot{\gamma}\eta_\text{p}}-\frac{\eta_\text{s}}{\eta_\text{p}}\right) \\
		0 &= 2\epsilon\lambda^2\dot{\gamma}^2\left[\partial_xp\frac{r}{2^j\dot{\gamma}\eta_\text{p}} - \frac{\eta_\text{s}}{\eta_\text{p}}\right]^2 + \ln\left(\partial_xp\frac{r}{2^j\dot{\gamma}\eta_\text{p}}-\frac{\eta_\text{s}}{\eta_\text{p}}\right)
		\intertext{This has the following form}
		0 &= ab^2 + \ln b
	\end{align}
	This can be solved using the Lambert W function as follows.
	\begin{align}
		b &= -\sqrt{\frac{W_0\left(2a\right)}{2a}}
		\intertext{Resolving for $r$ yields the following.}
		r &=  \frac{2^j\eta_\text{s}}{\partial_xp}\dot{\gamma} - \frac{\eta_\text{p}2^{j-1}}{\lambda\sqrt{\epsilon}\partial_xp}\sqrt{W_0\left(4\epsilon\lambda^2\dot{\gamma}^2\right)}
	\end{align}
	We notice, that this is the solution for a poiseuille flow with viscosity $\eta_\text{s}$ expanded by a term that applies a nonuniform scaling of the profile  along the $r$-axis towards higher radii. We simplify again using $c < 0$ and $d > 0$.
	\begin{equation}
		r = c\dot{\gamma} - R_\eta\frac{c}{d}\sqrt{W_0\left(d^2\dot{\gamma}^2\right)}
	\end{equation}
	Finally this function needs to be inverted. This needs to be done numerically. We calculate a range of $\dot{\gamma}$ from $0$ to $\frac{R}{c}$ with a small step interval. Here $R$ is the radius of the channel or pipe. This is guaranteed to contain all relevant radii. We flip the axis and integrate numerically using scipy.integrate.cumulative\_trapezoid\todo{cite}. This is very accurate as we conveniently picked the spacing of our now $x$-axis in a way, that the $y$-axis ($\dot{\gamma}$) varies little for each step. From the recovered curve $u$ is calculated at each required $r$ through linear interpolation.
